---
title: 'Part 1: Simulation Exercise Instructionsless'
author: "James Thomson"
date: "September 27, 2017"
output:
  html_document: default
  pdf_document: default
---

# Overview

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

# Simulations

Set the seed and needed parameters. Create a distribution for 1000 simulations of
40 exponentials.
```{r}
# Set seed value

set.seed(2345)

# Set parameters

sims <- 1000
n <- 40
lambda <- 0.2

# Run simmulation

mns = NULL
for(i in 1 : sims) {
        x <- rexp(n, lambda)
        mns <- c(mns, mean(x))
}
```

# Sample Mean versus Theoretical Mean

### Sample Mean

```{r}
smean <- mean(mns)
smean
```

### Theoretical Mean

```{r}
tmean <- 1/lambda
tmean
```
There is a very small difference between the Sample and Theoretical Means

```{r}
diffmean <- tmean - smean
diffmean
```

Plot a histogram of means

```{r}
hist(mns, breaks = 40, main = 'Distributions', xlab = 'Means')
curve(dnorm(x, mean = smean, sd = sd(mns))*100, col = "red", lwd = 2, add = TRUE)
```

The means are very close to a normal distribution

# Sample Variance verus Theoretical Variance

Sample Variance

```{r}
svar <- var(mns)
svar
```

Theoretical Variance

```{r}
tvar <- (1/lambda^2)/n
tvar
```
There is a very small difference between the Sample and Theoretical Variance

```{r}
diffvar <- tvar - svar
diffvar
```
Plot a histogram of variances

```{r}
hist(svar, breaks = 40, main = 'Distributions', xlab = 'Variance')
curve(dnorm(x, mean = svar, sd = sd(mns))*100, col = "red", lwd = 2, add = TRUE)
```

# Distribution

The Sample distribution is very close to the normal distribution

```{r}
qqnorm(mns); qqline(mns)
```
